---
title: "Introduction to SPS"
output: 
  md_document:
    variant: markdown_github
---

<br>

<p class = "h5">**On this page, we provide the introduction to `spsR` package (10 minute read).**</p> 

<br>
<div class="card bg-light mb-3">
  <div class="card-header">**Table of Contents**</div>
  <div class="card-body">
  1. [**Overview**](#step): <br> 
  We provide the overview of `spsR` package.

  2. [**Example: Multi-Country Survey Experiment**](#example): <br>
  We introduce an illustraitve example and its data set. 

  3. [**`sps`: Site Selection**](#sps): <br>
  We explain how to use function `sps()` for site selection

  4. [**`sps_estimator`: Estimate the Average-Site ATE**](#sps_estimator): <br>
  We explain how to use function `sps_estimator()` to analyze multi-site studies.
  
  </div>
</div>

<br>

### <span style="color:#3366CC">**1. \ Overview**</span> {#over}
<p class="lh-lg">
</p>
External validity of causal findings is a focus of long-standing debates. Whether and how can researchers generalize causal findings across different contexts and settings? To address this question, an increasing number of researchers use **a multi-site/multi-context design** where researchers conduct causal studies in multiple sites to compare and aggregate findings across contexts. 

When designing multi-site experiments (or observational studies), the fundamental research design question is **how should we select study sites for external validity?** 

The R package `spsR` aims to help researchers **select sites for external validity with statistical foundation, while accommodating practical and logistical constraints**. 

In particular, `spsR` implements Synthetic Purposive Sampling (SPS) introduced in [Egami and Lee (2023+)](https://naokiegami.com/paper/sps.pdf), which improves upon conventional purposive sampling by combining ideas from the synthetic control method --- it selects diverse sites such that non-selected sites are well approximated by the weighted average of the selected sites. By doing so, even without random sampling, we can make the weighted average of selected sites representative of all the sites, including non-selected sites in the population of sites. Please read [Egami and Lee (2023+)](https://naokiegami.com/paper/sps.pdf) for the methodological details. 

Using `spsR`, researchers can 

* Select diverse sites based on the SPS framework 
* Estimate the average causal effect for the population of sites 

<!-- On this page, we introduce the basic functionalities that is likely to be sufficient for most users, most of the time. For users who need more advanced functionalities, we offer AAA.  -->

<br>

### <span style="color:#3366CC">**2. \ Example: Multi-Country Survey Experiment**</span>  {#example}
<p class="lh-lg">
</p>
We use a multi-country survey experiment on attitudes toward immigrants (Naumann et al., 2018) as an example here. This study aims to answer a long-standing question in the immigration literature---whether and how much do natives prefer highly skilled migrants to low skilled migrants? To do so, in each site, they conduct a survey experiment where they randomly varied the skill level of hypothetical immigrant groups, i.e., “professionals” or “unskilled labors” (treatment variable), and asked respondents to report the support level for this immigrant group (outcome variable). 

```{r, eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(spsR)
data("X_Imm")
```

In this example, we try to select 6 countries from 15 European countries as study sites for the multi-site survey experiment. 

When selecting study sites for multi-site experiments, we choose site-level variables that we want to diversify. For example, to understand whether causal findings vary across different contexts, researchers might include diverse countries with different GDP, size of migrant population, unemployment rates, and so on. More formally, researchers should choose site-level variables that predict across-site heterogeneity of causal effects. 

In this illustration, we consider 8 variables discussed in the original paper: GDP (`GDP`), unemployment rates (`Unemployment`), size of migrant population (`Immigration`), the proportion of females (`Female`), the mean age (`Age`), the mean education (`Education`), general support for immigration (`Immig_Support`), and sub-regions in Europe (`Eastern Europe`, `Northern Europe`, `Southern Europe`, `Western Europe`). 

```{r, eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
colnames(X_Imm)
```

In practice, it is recommended to standardize each continuous variable to make the scale of site-level variables comparable. That is, each variable has the mean zero and the standard deviation one.  
```{r}
var_cont <-  c('GDP', 'Unemployment', 'Immigration', 
               'Female', 'Age', 'Education', 'Immig_Support')
X_Imm[, var_cont] <- scale(X_Imm[, var_cont])
round(head(X_Imm), 3)
```

Users can use `sps_plot()` to visualize the distribution of site-level variables before site selection. We select the first several variables that are continuous. 

```{r, fig.dim = c(7, 7), fig.align = 'center'}
sps_plot(X = X_Imm, before_selection = TRUE, columns = var_cont)
```
<br>

### <span style="color:#3366CC">**3. \ `sps`: Site Selection**</span> {#sps}
<p class="lh-lg">
</p>
We can use `sps()` to select 6 sites from 15 European countries (the population of sites). We start with the basic version of `sps` and illustrate refinements next. 

```{r}
out <- sps(X = X_Imm, N_s = 6)
```

**Arguments**

* `X`: Site-level variables for the population of sites. Row names should be names of sites.
* `N_s`: Number of study sites to be selected.

In this example, `sps` selected the following 6 sites. 

```{r}
out$selected_sites
```

Users can investigate the selected sites visually using `sps_plot()`. 

```{r, fig.dim = c(7, 7), fig.align = 'center'}
sps_plot(out, columns = var_cont)
```

Red triangles represent selected sites and black circles represent non-selected sites. In the last row and the diagonal plots, users can see the marginal distribution of each variable. All the remaining figures in the middle show bivariate relationships between two variables. 

We can see that selected sites successfully cover a wide range of values in each site-level variable. 

If necessary, users can refine site selection using *stratification* to make sure selected sites satisfy certain conditions. This stratification can help users to incorporate various practical, logistical, and ethical constraints. See some examples below. 

<details>
<summary><span style="color:#3366CC"><tag class = "h4">**Stratify SPS**</tag></span></font></summary>

In practice, researchers might want to stratify some variables to make sure that study sites satisfy certain conditions. 

#### <span style="color:#3366CC">Example 1: Improve Diversity of Site-Level Variables</span>
<p class="lh-lg">
</p>
It is often useful to select sites that have high, middle, low values in some site-level variables. For example, users might want to make sure that we select at least one country from low, medium, and high GDP countries. 

Users can enforce this condition using `stratify_sps().` 

To select countries with high GDP, we might want to select `at least` `2` sites that have `GDP` `larger than or equal to` `0.5` (i.e., 0.5 standard deviation because we standardized above). We can incorporate this information as follows. 

```{r}
st_GDP_1 <- stratify_sps(X = X_Imm, 
                         num_site=list("at least", 2), 
                         condition=list("GDP", "larger than or equal to", 0.5))
```

To select countries with medium GDP, we might want to select `at least` `2` sites that have `GDP` `between` `c(-0.25, 0.25)` (i.e., between -0.25 and 0.25 standard deviation). 

```{r}
st_GDP_2 <- stratify_sps(X = X_Imm, 
                         num_site=list("at least", 2), 
                         condition=list("GDP", "between", c(-0.25, 0.25)))
```

Finally, to select countries with low GDP, we might want to select `at least` `2` sites that have `GDP` `smaller than or equal to` `-0.5` (i.e., -0.5 standard deviation). 

```{r}
st_GDP_3 <- stratify_sps(X = X_Imm, 
                         num_site=list("at least", 2), 
                         condition=list("GDP", "smaller than or equal to", -0.5))
```

**Arguments**

* `num_site`: A list of two elements, e.g., `list("at least", 1)`. This argument specifies the number of sites that should satisfy `condition` specified below. The first element should be either `at least` or `at most`. The second element is integer. For example, `list("at least", 1)` means that we stratify SPS such that we select *at least 1* site that satisfies `condition` (specified below).

* `condition`: A list of three elements, e.g., `list("GDP", "larger than or equal to", 1)`. This argument specifies conditions for stratification. The first element should be a name of a site-level variable. The second element should be either `larger than or equal to`, `smaller than or equal to`, or `between`. The third element is a vector of length 1 or 2. When the second element is `between`, the third element should be a vector of two values. For example, `list("GDP", "larger than or equal to", 1)` means that we stratify SPS such that we select `num_site` sites that have *GDP larger than or equal to 1*.

After specifying these statification, researchers can supply them to `sps()`. 

```{r}
out_GDP <- sps(X = X_Imm, 
               N_s = 6, 
               stratify = list(st_GDP_1, st_GDP_2, st_GDP_3))
```

**Arguments**

* `stratify`: A `list` where each element is an output from function `stratify_sps()`.

```{r}
out_GDP$selected_sites
```


#### <span style="color:#3366CC">Example 2: Select Countries from Each Sub-Region</span>
<p class="lh-lg">
</p>
It is often useful to select sites from different subgroups. For example, users might want to make sure that we select at least one country from each of 4 sub-regions of Europe. 

The following code ensures to select `at least` `1` site from countries that have `Northern Europe = 1`.

```{r}
st_NE <- 
  stratify_sps(X=X_Imm, 
               num_site=list("at least", 1), 
               condition=list("Northern Europe", "larger than or equal to", 1))
```

Similarly, the following codes ensure that we select `at least` `1` site from countries that have `Eastern Europe = 1`, `Western Europe = 1` and `Southern Europe = 1`, respectively.

```{r}
st_EE <- 
  stratify_sps(X=X_Imm, 
               num_site=list("at least", 1), 
               condition=list("Eastern Europe", "larger than or equal to", 1))

st_WE <- 
  stratify_sps(X=X_Imm, 
               num_site=list("at least", 1), 
               condition=list("Western Europe", "larger than or equal to", 1))

st_SE <- 
  stratify_sps(X=X_Imm, 
               num_site=list("at least", 1), 
               condition=list("Southern Europe", "larger than or equal to", 1))
```

Finally, users combine them into one `list` and supply it to function `sps()`. 

```{r}
out_region <- sps(X = X_Imm, 
                  N_s = 6, 
                  stratify = list(st_NE, st_EE, st_WE, st_SE))
```

```{r}
out_region$selected_sites
```

Please see [Stratified SPS](http://naokiegami.com/spsR/articles/stratify_sps.html) for practical recommendations about stratification.

<!-- We can also visualize it.  -->

<!-- ```{r, fig.dim = c(7, 7), fig.align = 'center'} -->
<!-- sps_plot(out_region) -->
<!-- ``` -->

</details>

### <span style="color:#3366CC">**4. \ `sps_estimator`: Estimate the Average-Site ATE**</span> {#sps_estimator}
<p class="lh-lg">
</p>
Once we complete a randomized experiment (or an internally valid observational study) in each site, we now aggregate evidece for external validity.  

`sps_estimator()` will aggregate estimates of the average treatment effects (ATEs) in each specific site in order to estimate the average-site ATE (the average of the ATEs in the population of sites). In our example, the average-site ATE is the average of the ATEs across 15 European countries, including countries that were not selected. 

We start with ATE estimates in the selected sites. We use experimental estimates in selected 6 sites from Naumann et al (2018). Users only need to obtain point estimates and standard errors of the average treatment effects (ATEs) in each site. As an example, we use selected sites from `out`. 

```{r}
data("exp_estimate_Imm_selected")
exp_estimate_Imm_selected
```

For external validity analysis, we then aggregate these site-specific ATEs in selected sites to estimate the average-site ATE. 

```{r}
sps_est <- sps_estimator(out = out, 
                         estimates_selected = exp_estimate_Imm_selected)

sps_est2 <- sps_estimator(X = out$internal$X, estimates_selected = exp_estimate_Imm_selected, 
                          selected_sites = sort(unique(out$selected_sites)))
summary(sps_est)
summary(sps_est2)
```

**Arguments** 

* `out`: Output from function `sps()`
* `estimates_selected`: `data.frame` with two columns: the first column represents estimates of the site-specific ATEs for the selected sites and the second column represents its corresponding standard error. The number of rows is equal to the number of the selected sites and `rownames(estimates_selected)` should be names of the selected sites.

We can check the results using `summary` function.
```{r}
summary(sps_est)
```

We can also visualize both site-specific ATEs and the average-site ATE together using `plot()`.
```{r}
plot(sps_est)
```


Finally, users can use site-level cross-validation to assess the influence of unobserved moderators. In particular, users can randomly choose half of the selected sites as if they were unobserved non-selected sites and predict the average ATE of those non-selected sites based on the remaining selected sites. The null hypothesis is that the predicted ATEs are the same as the observed ATEs. 

```{r}
sps_cv_out <- sps_cv(out = out, estimates_selected = exp_estimate_Imm_selected)
round(sps_cv_out$p_value, 2)
```

In this example, an estimated p-value is `r round(sps_cv_out$p_value, 2)`, and we do not find evidence for the bias due to unobserved moderators. 
