---
title: "Handle Missing Data"
output: 
  md_document:
    variant: markdown_github
always_allow_html: yes
---

<br>

<p class="lh-lg">

**On this page, we describe how to use function `impute_var()` in R package `spsRdata` to impute missing values in a data set.**

<br>
**Note**: We offer function `impute_var`, which is simply a wrapper for two popular missing-data-imputation methods, `amelia` and `miceRanger`. Our goal is to facilitate data collection and data cleaning such that users can use functions in R package `spsR` easily. Please read papers and user guides of a missing-data-imputation method you choose ([Amelia](https://gking.harvard.edu/amelia) and [MICERanger](https://github.com/FarrellDay/miceRanger)).
</p>

<br>

### <span style="color:#3366CC">**Example: Impute Missing Values with `impute_var`**</span>
<p class="lh-lg">
</p>

```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(spsRdata)
```

`impute_var` in R package `spsRdata` imputes missing values in both continuous and categorical variables. 

Categorical variables (`character` or `factor`) are converted into dummies for each unique value prior to imputation. The data set can be either cross-sectional or time-series cross-sectional. Here, we demonstrate the utility of `impute_var` using the country-level data set `sps_country_data`, but `impute_var` can handle any type of data set.

```{r}
data(sps_country_data)
```

As an example, we will impute two variables, Polity score (`e_p_polity`), GDP (`NY.GDP.MKTP.CD`), and social group exclusion index (`v2xpe_exlsocgr`), which have some missing values. 

```{r}
apply(sps_country_data[, c('e_p_polity', 'NY.GDP.MKTP.CD', 'v2xpe_exlsocgr')], 2, function(x) sum(is.na(x)))
```

We impute the two variables using `amelia`. By specifying `id_site` and `id_time`, function `impute_var` can respect the time-series cross-sectional data structure. When we only specify `id_site`, the underlying imputation method assumes the cross-sectional data. 

```{r, eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE, results = FALSE}
imputed_data <- impute_var(data = sps_country_data, 
                           id_unit = 'country', 
                           id_time = 'year', 
                           var_impute = c('e_p_polity', 'NY.GDP.MKTP.CD'),
                           var_ord = 'e_p_polity',
                           var_lgstc = 'v2xpe_exlsocgr',
                           method = "amelia")
```

```{r}
apply(imputed_data[, c('e_p_polity', 'NY.GDP.MKTP.CD', 'v2xpe_exlsocgr')], 2, function(x) sum(is.na(x)))
```

**Arguments**

* `data`: A `data.frame` containing variables to impute.
* `id_site`: A unique identifier for sites. A column name in `data`. 
* `id_time`: A unique identifier for time index. A column name in `data`. If unspecified (`NULL`; default), it assumes `data` is cross-sectional.
* `var_impute`: A vector with one or more variable names for which imputation is performed. If unspecified (`NULL`), it imputes all variables in `data` except for the `id_site` and `id_time` variables.
* `var_ord`: A vector with one or more variable names that is ordinal for which imputation is performed. Binary variable names can be included in eithwe `var_ord` or `var_nom`. 
* `var_nom`: A vector with one or more variable names that is nominal for which imputation is performed. 
* `var_lgstc`: A vector with one or more variable names that is proportional ranging between 0 and 1 for which imputation is performed. 
* `var_predictor`: A vector with one or more variable names that we use as predictors to impute variables in `var_impute`. If unspecified (`NULL`), the function uses all variables in `data` except for variables in `var_impute`.
* `method`: Imputation method. Choose `amelia`, `mice` or `miceranger`. Default is `amelia`. 
* `n_impute`: The number of imputed data sets to produce (equivalent of argument `m` in `amelia()` and `miceRanger()`). Default is 5.
* `...`: Arguments passed onto `Amelia::amelia()` or `mice::mice()` or `miceRanger::miceRanger()`.

**Imputation Methods**

1. [Amelia](https://gking.harvard.edu/amelia). Please read Honaker, King, and Blackwell (2011) for more details.
2. [MICE](https://amices.org/mice/). Please read Van Buuren and Groothuis-Oudshoorn (2011) for more details.
3. [MICERanger](https://github.com/FarrellDay/miceRanger). Please read Wilson (2020) for more details.


#### **Practical Suggestions**

Missing data imputation works better when users supply more variables and observations because the underlying method can learn relationships between different variables better. Therefore, in practice, we recommend imputing missing data first before sub-setting the data to the target population of sites and site-level variables they diversify. For example, if users want to use `sps_country_data` we provide in `spsRdata`, they can impute missing data first using the full data and then subset the data to focus on their target population of sites and site-level variables of interest. 

<br>

### References

Honaker, J., King, G., & Blackwell, M. (2011). Amelia II: A program for missing data. Journal of statistical software, 45, 1-47.

King, G., Honaker, J., Joseph, A., & Scheve, K. (2001). Analyzing incomplete political science data: An alternative algorithm for multiple imputation. American political science review, 95(1), 49-69.

Van Buuren, S., & Groothuis-Oudshoorn, K. (2011). mice: Multivariate imputation by chained equations in R. Journal of statistical software, 45, 1-67.

Wilson, S. (2020). miceRanger: Multiple imputation by chained equations with random forests. R package version, 1(5).
