---
title: "Prepare Country-Level Variables"
output: 
  md_document:
    variant: markdown_github
always_allow_html: yes
---

<br>

<p class = "h5">**On this page, we explain how to prepare country-level variables with `spsRdata` package (10 minute read).**</p> 

<br>
<div class="card bg-light mb-3">
  <div class="card-header">**Table of Contents**</div>
  <div class="card-body">
  1. [**Overview**](#over): <br> 
  We provide the overview of `spsRdata` package.

  2. [**Default Country-Level Data set**](#default): <br>
  We introduce a default country-level data set to ease data collection.

  3. [**Prepare Your Own Country-Level Data**](#own): <br>
  We explain how to prepare your own country-level data using several functions in `spsRdata`
     
     * [Start with `sps_country_data`](#start) 
     
     * [Merge Country-Level Variables from V-Dem or World Bank Data Sets](#v-dem)
     
     * [Merge External Data sets](#external) 
     
     * [Check Data](#summary) 
     
     * [Clean Data for `sps()`](#clean) 
  
  </div>
</div>

<br>

### <span style="color:#3366CC">**1. \ Overview**</span> {#over}
<p class="lh-lg">
</p>
The purpose of R package `spsRdata` is to facilitate data collection for **multi-country** studies. Users can supply data they prepared on this page in function `sps()` for site selection.

Users can start with the default data set `sps_country_data` that contains a wide range of commonly-used country-level variables in the social sciences. R package `spsRdata` also offers several functions to merge additional country-level variables from well-known data sets (e.g., V-Dem and World Bank) as well as from any external data users want to add. 

*Note*: We emphasize that R package `sps` and the corresponding SPS algorithm can be applied to any type of study site (e.g., states, districts, villages, or schools). As *multi-country* studies are most popular in political science, R package `spsRdata` focuses on country-level data collection for now, but we are planning to add more functionalities to facilitate data collection at different levels (e.g., states and cities) in the US as well. 

### <span style="color:#3366CC">**2. \ Default Country-Level Data set** </span> {#default}
<p class="lh-lg">
</p>
```{r eval = TRUE, echo = TRUE, tidy=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(spsRdata)
```

In R package `spsRdata`, we provide a country-level data set (`sps_country_data`) comprised of 22 variables for 179 countries between 2010 and 2022. It contains a wide range of variables commonly used in the social sciences: they represent political, economic, and social characteristics of countries. These variables are collected from the following sources. 

* [Varieties of Democracy (V-Dem) project](https://v-dem.net/data/the-v-dem-dataset)
    + V-Dem data set collects variables from a variety of sources including United Nations, Quality of Government, Clio Infra, and Polity5 Project, and currently includes the most comprehensive political indicators. 
        
* [World Bank Indicators](https://datahelpdesk.worldbank.org/knowledgebase/articles/889392-about-the-indicators-api-documentation)
* [UN Statistics](https://unstats.un.org/unsd/methodology/m49/overview)
* [University of Groningen's World Languages](https://resourcewatch.org/data/explore/soc_071_world_languages?section=Discover&selectedCollection=&zoom=3&lat=0&lng=0&pitch=0&bearing=0&basemap=dark&labels=light&layers=%255B%257B%2522dataset%2522%253A%252220662342-dcdd-4a42-9f58-bcc80217de71%2522%252C%2522opacity%2522%253A1%252C%2522layer%2522%253A%2522f2d76e6b-060d-4dc9-83ea-284bef6b2aae%2522%257D%255D&aoi=&page=1&sort=most-viewed&sortDirection=-1)

`sps_country_data` includes the following variables.

```{r}
data(sps_country_data)
head(sps_country_data, n = 3)
```
**Descriptions of Variables:**

```{r, echo = F}
if(!require(kableExtra)) install.packages("kableExtra")
if(!require(knitr)) install.packages("knitr")

library(kableExtra)
library(knitr)
kbl(sps_country_codebook) %>%
  kable_classic_2(full_width = T) %>%
  kable_styling(font_size = 12, bootstrap_options = c('hover', 'condensed')) %>%
  column_spec(1, italic = T) %>%
  pack_rows('Key Identifiers', 1, 3) %>%
  pack_rows('Geographic Indicators', 4, 6) %>%
  pack_rows('Democracy Indicators', 7, 9) %>%
  pack_rows('Political Indicators', 10, 15) %>%
  pack_rows('Economic Indicators', 16, 17) %>%
  pack_rows('Sociodemographic Indicators', 18, 20) %>%
  pack_rows('Social Indicators', 21, 22)
```

Users can check the codebook using `data(sps_country_codebook)`. 

```{r}
data(sps_country_codebook)
sps_country_codebook[sps_country_codebook$Variable %in% 
                       c("country", "year", "NY.GDP.MKTP.CD", "SP.POP.TOTL"), ]
```


<br>


### <span style="color:#3366CC">**3. \ Prepare Your Own Country-Level Data**</span> {#own}
<p class="lh-lg">
</p>
We follow our example in the [Get Started Page](http://naokiegami.com/spsR/articles/intro.html) and use a multi-country survey experiment by Naumann et al. (2018) to illustrate how to use `sps_country_data` and various functions in `spsRdata`. 

#### <span style="color:#3366CC">**3.1 \ Start with `sps_country_data`**</span> {#start}
<p class="lh-lg">
</p>
In our example, we are interested in 8 country-level variables discussed in the original paper: GDP, unemployment rates, size of migrant population, general support for immigration, the proportion of female, the mean age, the mean education, and sub-regions in Europe. We illustrate how to collect these variables step by step. 

We are interested in 15 European countries and variables measured in 2015 (the year in which Naumann et al (2018)'s experiments were conducted). If users are running experiments now, they can, of course, choose site-level variables measured in most recent years. 

```{r}
country_use <- c("Austria", "Belgium", "Czechia", "Denmark", "Finland", 
                 "France", "Germany", "Ireland","Netherlands", "Norway", 
                 "Slovenia", "Spain", "Sweden", "Switzerland", "United Kingdom")

sps_country_data1 <- 
  sps_country_data[sps_country_data$year == 2015 & 
                     sps_country_data$country %in% country_use,]
```

Three out of the 8 variables are already in our default data set: sub-regions (`subregion`), GDP (`NY.GDP.MKTP.CD`) and unemployment rate (`SL.UNEM.TOTL.ZS`). So we subset the data set to include only those variables we're interested in. 

```{r}
sps_country_data2 <- sps_country_data1[, c("country", "iso3", 
                                           "subregion", "year", 
                                           "NY.GDP.MKTP.CD", "SL.UEM.TOTL.ZS")]
```

<br>

#### <span style="color:#3366CC">**3.2 \ Merge Country-Level Variables from V-Dem or World Bank Data Sets**</span> {#v-dem}
<p class="lh-lg">
</p>
Users can also search for other country-level variables in V-Dem and World Bank using keywords via the function ``search_var``. Here, we look for data on size of migrant population:

```{r}
search_result <- search_var(keyword = "international migrant")
head(search_result, n = 3)
```

Based on the search return, users can add variable(s) to the baseline data set using ``merge_data``. In our example, we're interested in adding `SM.POP.TOTL.ZS` that captures the proportion of migrant population.

```{r}
sps_country_data3 <- merge_data(data = sps_country_data2, 
                                vars = c("SM.POP.TOTL.ZS"))
```
**Arguments**

* `data`: A `data.frame` where the new variables should be merged onto. 
* `vars`: A vector with one or more variable names that should be merged onto `data`. We merge variables from V-Dem or World Bank data sets (use ``search_var()`` to search for names of new variables).

```{r}
head(sps_country_data3, n = 3)
```


<br>


#### <span style="color:#3366CC">**3.3 \ Merge External Data sets**</span> {#external}
<p class="lh-lg">
</p>
Users can also merge their own external data set. In our example, 4 variables---the proportion of female, the mean age, the mean education, and general support for immigration---come from European Social Survey data. As an illustration, we included this external data as `ess_data`.

```{r}
data(ess_data)
head(ess_data, n = 3)
```
First, it is strongly recommended to standardize country names using ISO3 code. Users can do this easily with R package `countrycode`.

```{r, echo = F}
if(!require(countrycode)) install.packages("countrycode")
```


```{r}
library(countrycode)
ess_data$iso3 <- countrycode(sourcevar = ess_data$country, 
                             origin = "country.name", destination = "iso3c")
```

Then, users can use the same function ``merge_data`` to merge their own external data set as well.

```{r}
sps_country_data4 <- merge_data(data = sps_country_data3, 
                                newdata = ess_data, id_site = 'iso3')
```

**Arguments**

* `data`: A `data.frame` where the new variables should be merged onto. 
* `newdata`: A `data.frame` that contains new variables to be merged on. When `newdata = NULL,` the function merges variables from V-Dem or World Bank data sets.
* `vars`: A vector with one or more variable names that should be merged onto `data`. When `newdata = NULL,` this must be specified  (use ``search_var()`` to search for new variables). When `newdata` is specified and `vars = NULL` (default), the function merges all variables in `newdata`.
* `id_site`: A unique identifier for sites. A column name in `data`. 
* `id_year`: A unique identifier for year. A column name in both `data` and `newdata`. If `NULL` (default), the function assumes `newdata` is the cross-sectional data and the same values in `newdata` will be merged to all years in `data`. Ignored if `newdata` is `NULL`.

<!-- The function uses fuzzy-match to match country names in two data sets when an exact match does not exist. Here, we discover that the two countries---Czech Republic and UK---are being matched incorrectly from this process. So we instead use an ISO3 code to re-merge the data: -->

<!-- ```{r} -->
<!-- library(countrycode) -->
<!-- ess_data$iso3 <- countrycode(sourcevar = ess_data$country, origin = "country.name", destination = "iso3c") -->
<!-- ess_data <- ess_data[,which(names(ess_data) != 'country')] -->
<!-- head(ess_data, n = 3) -->
<!-- sps_country_data4 <- merge_data(data = sps_country_data3, newdata = ess_data, id_site = 'iso3') -->
<!-- head(sps_country_data4[sps_country_data4$country == 'Czechia',]) -->
<!-- ``` -->

#### <span style="color:#3366CC">**3.4 \ Check Data**</span> {#summary}
<p class="lh-lg">
</p>
Users can run simple descriptive analyses to examine distribution and missing values. The function ``desc_stat`` returns simple descriptive statistics and number of missing entries in each year in the data set. Character and factor variables are automatically binarized to obtain descriptive statistics:

```{r}
desc <- desc_stat(data = sps_country_data4)
head(desc$desc, n = 3)
head(desc$miss, n = 3)
```
A list object `desc` returns descriptive summary of each variable in the data set, whereas `miss` returns total counts of missing values (i.e., number of missing countries) in each year in the data set. If the years in which a user is interested in using contains a lot of missing values, we suggest imputing missing values using the `impute_var` function in R package `spsR`. Please see [Handle Missing Data](http://naokiegami.com/spsR/articles/missing_data.html).

#### <span style="color:#3366CC">**3.5 \ Clean Data for `sps()`**</span> {#clean}
<p class="lh-lg">
</p>
Finally, we clean data for function `sps()`. First, we need to have informative row names for the data. 

```{r}
row.names(sps_country_data4) <- sps_country_data4$country
```

Second, for categorical variables, we binarize them. 

```{r}
X_region <- model.matrix(~ subregion - 1, 
                         data = sps_country_data4[, "subregion", drop = FALSE])

colnames(X_region) <- c("Eastern Europe", "Northern Europe", 
                        "Southern Europe", "Western Europe")
```

Third, we keep all relevant variables and rename them. 

```{r}
X_Imm <- cbind(sps_country_data4[, c("NY.GDP.MKTP.CD", "SL.UEM.TOTL.ZS", 
                                     "SM.POP.TOTL.ZS", "female", "age", 
                                     "education", "support_imm")], 
               X_region)

colnames(X_Imm) <- c('GDP', 'Unemployment', 'Immigration', 'Female', 'Age', 
                     'Education', 'Immig_Support', 
                     colnames(X_region))
head(X_Imm, n = 3)
```

This `X_Imm` is ready for function `sps()`.

<!-- ```{r} -->
<!-- round(desc$corr[1:3, 1:3], 3) -->
<!-- ``` -->

<!-- A list object `corr` returns correlation matrix for all pairs of variables in the dataset. Highly correlated variables mean that they are capturing similar cross-country contexts, so users might want to consider dropping one of them prior to implementing SPS.  -->

<br>

# References

Michael Coppedge et al. 2023. “V-Dem Dataset V13.” Varieties of Democracy (V-Dem) Project. https://doi.org/10.23696/VDEMDS23.

University of Groningen. 2016. “World Languages.” University of Groningen Open Data. www.resourcewatch.org.
